### Repr Res Peer Assessment 1

Richard Anderson


#### Loading and preprocessing the data

```{r, echo=TRUE}
d <- read.csv("activity.csv",header=TRUE)
```

#### What is the mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day 

```{r, echo=TRUE}
s <- split(d,d$date)
```

```{r, echo=TRUE,fig.width=4,fig.height=4}
hist(sapply(s, function(x) sum(x[,1],na.rm=TRUE)),main="Total number of steps taken each day")
```

2. Calculate and report the mean and median
number of steps taken each day.

```{r, echo=TRUE}
mean(sapply(s, function(x) sum(x[,1],na.rm=TRUE)))
median(sapply(s, function(x) sum(x[,1],na.rm=TRUE)))
```

#### Average daily activity pattern

1. Make a time series of the 5 min interval (x-axis) and the average number
of steps taken, avgd across all days (y-axis).

```{r, echo=TRUE, fig.width=4,fig.height=4}
s <- split(d,d$interval)
interval <- d[d$date=="2012-10-01","interval"]
activity <- sapply(s, function(x) mean(x[,1],na.rm=TRUE)) #Interval means
plot(interval,activity,type="l",ylab="Number of steps")
```

2. Which 5 min interval, on average across all days, contains the
maximum number of steps?

```{r, echo=TRUE}
names(which.max(activity))
```


#### Inputting missing values

1. Calculate and report total number of missing values in dataset.

```{r, echo=TRUE, fig.width=6,fig.height=4}
sum(!is.na(d$steps))    #number of missing values (rows w/NAs)
```

2. Devise strategy for filling in missing values in the dataset.

 I will use the mean for each 5 minute interval.

```{r, echo=TRUE}
fill <- rep(activity,61)
steps.fill <- numeric(length(d$steps))
#This loop fills in NAs with mean for each interval
for (i in 1:length(d$steps)) {
        if (is.na(d$steps[i])) {
                steps.fill[i] <- fill[i]
        } else {
                steps.fill[i] <- d$steps[i]
        }
}
```

3. Create a new dataset equal to the original but with NAs filled in.

```{r, echo=TRUE}
d.fill <- data.frame(d,steps.fill) #new filled data.frame
head(d.fill)
```

4. Make a histogram of the new dataset and report the new mean and median.

```{r, echo=TRUE, fig.width=4,fig.height=4}
s <- split(d.fill,d.fill$date)
hist(sapply(s, function(x) sum(x[,4],na.rm=TRUE)),main="Total number of (imputed) steps taken each day")
mean(sapply(s, function(x) sum(x[,4],na.rm=TRUE)))
median(sapply(s, function(x) sum(x[,4],na.rm=TRUE)))
```

#### Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor weekday vs. weekend factor variable.

```{r, echo=TRUE}
#Create factor weekday vs. weekend variable
wknd <- character(length(d.fill$steps))
for (i in 1:length(d.fill$steps)) {
        if (weekdays(as.Date(d.fill$date[i]))=="Saturday" |
            weekdays(as.Date(d.fill$date[i]))=="Sunday") {
                wknd[i] <- "weekend"
        } else {
                wknd[i] <- "weekday"
        }
}
d.fill.wknd <- data.frame(d.fill,as.factor(wknd))
str(d.fill.wknd)
```

2. Create panel plot containing time series plot of average number
of steps for weekdays vs. weekends.

```{r, echo=TRUE, fig.width=4,fig.height=4}
#Create weekday vs. weekend activity pattern means
s <- split(d.fill.wknd,d.fill.wknd$interval)
activity.fill.wkdy <- sapply(s, function(x) 
        mean(x[x$as.factor.wknd.=="weekday",4]))
activity.fill.wknd <- sapply(s, function(x) 
        mean(x[x$as.factor.wknd.=="weekend",4]))
#Create data frame with weekday vs. weekend activity pattern mean data
weekend <- gl(2,288,labels=c("weekday","weekend"))
activity.fill.spl <- c(activity.fill.wkdy,activity.fill.wknd)
activity.wk <- data.frame(activity.fill.spl,rep(interval,2),weekend)
names(activity.wk) <- c("activity","interval","weekend")
str(activity.wk)
#Create panel plot using lattice
library(lattice)
xyplot(activity ~ interval | weekend, data = activity.wk,type="l",
       layout=c(1,2),ylab="Number of steps",xlab="Interval")
```








